GPU-friendly Stroke Expansion

### 2024-06-14 paper.tex
Edits around dashing
- More explicitly state that we did not fully explore the solution
  space.
- Remove repeated text an the introduction.

Stroke rendering is an essential and widely part of the standard vector graphics imaging model. The standard representation of paths is a sequence of cubic Bézier segments. While there are a number of published techniques for GPU rendering of filled paths, stroke rendering is more challenging to parallelize, largely because path segments cannot be processed independently of each other; rendering of the \emph{joins} between path segments depends on both adjoining segments, and the endpoints of paths (when not closed) are rendered with \emph{caps}. Joins and caps can be rendered with a variety of styles within the same document. The style of stroke rendering also optionally includes \emph{dashing.}

@@

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments and the \emph{inner contours of joins} are required in addition to the parallel curves. The technique described in this paper produces strongly correct outlines according to this definition.

@@

Regarding dashing, previous work, particularly \citet{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages. Whether moving dashing to a GPU stage is a performance win depends on the scene and other factors in the overall rendering pipeline. While we would expect further performance improvement from GPU-based dashing, we were able to achieve acceptable performance with dash processing on CPU.

Previous work, particularly \citet{Kilgard2020}, has explained how to process dashing patterns on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages, and we were able to achieve acceptable performance with the dash processing on CPU. Thus, whether moving dashing to a GPU stage is a performance win depends on the scene and other factors in the overall rendering pipeline. Even so, for scenes that have a very large number of dashed strokes, we would expect further performance improvement from GPU-based dashing.

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel. The CPU encoding times (prior to the GPU prefix sums) for some of the large scenes is shown in Table~\ref{table:bump-estimate-cost}. Our implementation applies dashing on the CPU. It is conceptually fairly straightforward to implement on GPU; the core algorithm is prefix sum applied to segment length, and the Euler spiral representation is particularly well suited. However, there are complex implementation tradeoffs, so the dash style applied to the \emph{long dash} scene had to be processed on the CPU and dashes were sequentially encoded as individual strokes. Including CPU-side dashing, the time to encode \emph{long dash} on the Apple M1 Max was approximately 25~ms on average, while the average time to encode the path after dashing was 4.14~ms.

>>>>

Stroke rendering is an essential and widely part of the standard vector graphics imaging model. The standard representation of paths is a sequence of cubic Bézier segments. While there are a number of published techniques for GPU rendering of filled paths, stroke rendering is more challenging to parallelize, largely because path segments cannot be processed independently of each other; rendering of the \emph{joins} between path segments depends on both adjoining segments, and the endpoints of paths (when not closed) are rendered with \emph{caps}. Joins and caps can be rendered with a variety of styles within the same document. The style of stroke rendering also optionally includes \emph{dashing}, which applies a pattern of dashes and gaps to the outline of the shape.

@@

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments and the \emph{inner contours of joins} are required in addition to the parallel curves. The technique described in this paper produces strongly correct outlines according to this definition. Examples of weakly and strongly correct rendering are shown in Figure~\ref{fig:weak_strong}.

\begin{figure}
    \includegraphics[scale=0.5]{weak_strong}
    \caption{Example of weakly correct (top) and strongly correct (bottom) stroke rendering. The left example has a curvature cusp and requires evolutes to render correctly, while the right requires the inner join.}
    \Description{Two examples of strokes. The top left is missing an evolute section, making it appear that a half-disc has been carved out. The top right is missing inner joins, looking like spokes rather than a connected path. The bottom two examples are rendered with the full strokes.}
    \label{fig:weak_strong}
\end{figure}

@@

Previous work, particularly \citet{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages. We anticipate that the potential performance benefit of implementing dashing on a GPU depends on the scene and other factors in the overall rendering pipeline. While we would expect further performance improvement from GPU-based dashing, we did not fully explore the solution space and we were able to achieve acceptable performance with dash processing on CPU.

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel. The CPU encoding times (prior to the GPU prefix sums) for some of the large scenes is shown in Table~\ref{table:bump-estimate-cost}. Our implementation applies dashing on the CPU. It is conceptually straightforward to implement on GPU; the core algorithm is prefix sum applied to segment length, and the Euler spiral representation is particularly well suited. However, we believe the design of an efficient GPU pipeline architecture involves comples tradeoffs which we did not sufficiently explore for this paper. Therefore the dash style applied to the \emph{long dash} scene was processed on the CPU and dashes were sequentially encoded as individual strokes. Including CPU-side dashing, the time to encode \emph{long dash} on the Apple M1 Max was approximately 25~ms on average, while the average time to encode the path after dashing was 4.14~ms.
